---
title: "Introduction and EDA for EDDB"
author: "Brian Richards"
output: 
  html_notebook:
    toc: true
    toc_float: true
---

The Elite: Dangerous Database (EDDB) is a collection of data for the game 
Elite: Dangerous. The data is split into several different databases: 

* Prices of goods (listings.csv)
* Stations (stations.json)
* Populated Systems (systems_populated.json)
* Factions (factions,json)
* Commodities (commodities.json)
* Modules (modules.json)

This notebook will be an introduction to each data set and provide some 
exploratory data analysis of the data sets. This EDA will also act as 
prototypes for the information I will want to display in the Shiny app.

```{r, message = FALSE}
library(tidyverse)
```

## Prices

The `listings.csv` file gives a read out of the prices of goods at the time 
the file was updated.

```{r, message = FALSE}
listings <- read_csv("data/listings.csv", progress = FALSE)
```

```{r}
listings
```

The columns for the `listings` are:

* id: The id of the particular good
* station_id: The station where the good is being sold
* commodity_id: The good being sold (related to commodities.json)
* supply: How much of the item is available for purchase
* supply_bracket: Unsure of purpose
* buy_price: Price the good can be bought at 
* sell_price: Price the good can be sold at
* demand: How many items the station wants
* demand_bracket: Unsure of purpose
* collected_at: Time the data was collected


