---
title: "Introduction and EDA for EDDB"
author: "Brian Richards"
output: 
  html_notebook:
    toc: true
    toc_float: true
---

## Introduction

The Elite: Dangerous Database (EDDB) is a collection of data for the game 
Elite: Dangerous. The data is split into several different databases: 

* Prices of goods (listings.csv)
* Stations (stations.json)
* Populated Systems (systems_populated.json)
* Factions (factions,json)
* Commodities (commodities.json)
* Modules (modules.json)

This notebook will be an introduction to each data set and provide some 
exploratory data analysis of the data sets. This EDA will also act as 
prototypes for the information I will want to display in the Shiny app.

```{r, message = FALSE}
library(tidyverse)
library(jsonlite)
library(plotly)
```

## Prices

The `listings.csv` file gives a read out of the prices of goods at the time 
the file was updated.

```{r, message = FALSE}
listings <- read_csv("data/listings.csv", progress = FALSE)
```

```{r}
head(listings)
```

The columns for the `listings` are:

* id: The id of the particular good
* station_id: The station where the good is being sold
* commodity_id: The good being sold (related to commodities.json)
* supply: How much of the item is available for purchase
* supply_bracket: Unsure of purpose
* buy_price: Price the good can be bought at 
* sell_price: Price the good can be sold at
* demand: How many items the station wants
* demand_bracket: Unsure of purpose
* collected_at: Time the data was collected

We can then look at a summary of the data to get an idea of what is in the 
data set.

```{r}
summary(listings)
```

From the summary, we can see that there are number of NA's in the `supply_bracket` 
and `demand_backet` columns. Since we don't know what those do, we can ignore 
those for the time being. Additionally, there are zeroes in the `sell_price` 
and `buy_price` columns. Those are essentially NA's as a `sell_price` of 0 means 
that you can't sell the product there and a `buy_price` of 0 means there are 
none available at that station.

```{r}
listings %>% 
  filter(commodity_id == 1) %>% 
  select(buy_price, sell_price) %>% 
  gather(type, price) %>% 
  filter(price > 0) %>% 
  group_by(type) %>% 
  mutate(average = mean(price), median = median(price)) %>% 
  ggplot(aes(x = price, fill = type)) +
  geom_histogram(bins = 100) +
  facet_grid(type ~.) +
  geom_vline(aes(xintercept = average)) +
  geom_vline(aes(xintercept = median), linetype = "dotted")
```

This figure gives a quick histogram of the buy and sell prices of a single 
commodity across all available stations. Note that it filters out any buy or 
sell prices of 0. The solid line is the mean of the distribution and the dotted 
line is the median of the distribution. From the figure, an interesting topic 
to investigate is looking at comparative buy and sell prices across stations 
to create a system to search for the max difference.

```{r}
listings %>% 
  filter(station_id == 12) %>% 
  select(buy_price, sell_price) %>% 
  gather(type, price) %>% 
  filter(price > 0) %>% 
  group_by(type) %>% 
  mutate(average = mean(price), median = median(price)) %>% 
  ggplot(aes(x = price, fill = type)) +
  geom_histogram(bins = 20) +
  facet_grid(type ~.) +
  geom_vline(aes(xintercept = average)) +
  geom_vline(aes(xintercept = median), linetype = "dotted")
```

This figure looks at a histogram of buy and sell prices for a specific station 
(across all available commodities). Again, the solid line is the mean and the 
dotted line is the median.

### Questions 

I want to put together some initial questions to answer using my Shiny app. 

* What is the biggest price difference?
* What is the distance between the stations with the biggest price difference? 
* Given a particular commodity, where can I find it and at what price?
* Given a particular commodity, where can I sell it and at what price?
* How do the buying and selling price compare the the galatic average (and 
reported average average and median)?

## Stations 

```{r}
stations <- as_tibble(fromJSON("data/stations.json"))
```

```{r}
head(stations)
```

The stations data has 39 columns in it, the most relevant of which are the 
following:

* id: The station's id
* name: The station's name
* system_id: The system id where the station resides
* updated_at: Time at which the data was updated
* distance_to_star: Distance of the station from the system's star
* has_ : A set of booleans showing what amenities the station offers
* is_planetary: A boolean showing whether the station is on a planet
* selling_ships: A list of the ships being sold

There are also a number of other useful variables showing additional information 
about the stations.

We can summarize the most relevant columns and get a sense of distributions and 
any possible NAs.

```{r}
stations %>% 
  select(id, system_id, updated_at, distance_to_star, starts_with("has_"), is_planetary) %>% 
  summary()
```

We can also extract a list of possible ships that are sold and from that we 
can determine what stations sell which ships.

```{r}
stations %>% 
  select(selling_ships) %>% 
  unnest() %>% 
  distinct()
```

```{r}
stations %>% 
  select(id, name, system_id, selling_ships) %>% 
  unnest() %>% 
  filter(selling_ships == "Imperial Cutter") %>% 
  head()
```

For instance, we can get a listing of all of the stations that sell the 
"Imperial Cutter" and what system that station is in. 

### Questions 

* What are the list of purchasable ships?
* What ship can be bought where?
* What module can be bought where?

## Populated Systems

The populated systems data gives information on systems in the universe that 
are populated.

```{r}
populated_systems <- as_tibble(fromJSON("data/systems_populated.json"))
```

```{r}
head(populated_systems)
```

Like the stations data, the populated systems data has a large number of 
columns (29) that are mostly metadata about the systems. Most of the metadata 
is centered on government types, alleigance, security level, power play info, 
and factions. The information that will likely be useful for us are the 
following columns: 

* id: The system's id
* name: the system's name
* x, y, z: The coordinates of the system in the universe

First, I want to look at the summary statistics for the coordinates to verify 
that there are no missing coordinates. It looks like all of the coordinates 
are present.

```{r}
populated_systems %>% 
  select(x, y, z) %>% 
  summary()
```

I can make a 3D scatter plot of the coordinates of the systems and show the 
color of the controlling power. All of this done with plotly and could be a fun 
way of showing where a particular system is in comparison to other systems. In 
this plot I've only included the top 4 controlling powers (based on number of 
systems controlled). Since I haven't used plotly much, this could be a great 
opportunity to become more familiar with it.

```{r}
p <- populated_systems %>%
  filter(!is.na(power)) %>% 
  group_by(power) %>%
  mutate(control_number = n()) %>%
  ungroup() %>% 
  mutate(control_number = dense_rank(desc(control_number))) %>% 
  filter(control_number < 5) %>%
  plot_ly(x = ~x, y = ~y, z = ~z, color = ~power) %>% 
  add_markers(opacity = 0.1)

p
```

### Questions 

* Where are the various systems?
* What powers or factions control each system?

